---
title: "Homework 2 Part 1. My own question and solution"
author: "Runan Yao"
date: "September 21, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(stringr)
```

## Question:

When our member travel between different states, which fast food/restaurant is most popular ? 

Note:

1. Member home location can be found in member profile. It is changing overtime. 
So, when matching transaction with member profile, we need match by both member Id and transaction date.

2. Restaurant and fast food having MCCcode: 5812 and 5814. These code are one column of transaction data.

3. Restaurant and fast food store location can be found in transcation data.

## Solution:

### 1. Load data from file:

```{r}
trans1 <- read.csv('Data/April Vintage Transactions page 1.csv', stringsAsFactors=FALSE)
trans2 <- read.csv('Data/April Vintage Transactions page 2.csv', stringsAsFactors=FALSE)
trans2$DebtDimId <- as.character(trans2$DebtDimId)
trans3 <- read.csv('Data/April Vintage Transactions page 3.csv', stringsAsFactors=FALSE)
trans3$DebtDimId <- as.character(trans3$DebtDimId)
trans4 <- read.csv('Data/April Vintage Transactions page 4.csv', stringsAsFactors=FALSE)
trans4$DebtDimId <- as.character(trans4$DebtDimId)

trans.dat <- bind_rows( bind_rows(trans1,trans2),bind_rows(trans3,trans4) )
nrow(trans.dat)
head(trans.dat)
summary(trans.dat)

profile1 <- read.csv('Data/April Vintage 2013.csv', stringsAsFactors = FALSE)
profile2 <- read.csv('Data/April Vintage 2013 _ page 2.csv', stringsAsFactors = FALSE)
prof.dat <- bind_rows(profile1, profile2)
nrow(prof.dat)
head(prof.dat)
summary(prof.dat)


rm(trans1, trans2, trans3, trans4, profile1, profile2)
```

### 2. Table 1: Id, Date(start, end), Home state

Replace NULL DatePrevStmt by FDROpenDt if it is the first record. 
Replace NULL DatePrevStmt by StatementDt - 30 if it is not the first record
```{r}
temp <- prof.dat
idx <- which(( is.na(temp$DatePrevStmt) | temp$DatePrevStmt == 'NULL' ) & temp$MOB == 0)
temp[idx,]$DatePrevStmt <- temp[idx,]$FDROpenDt

idx <- which(is.na(temp$DatePrevStmt) | temp$DatePrevStmt == 'NULL')
temp[idx,]$DatePrevStmt <- as.character( as.Date(temp[idx,]$StatementDt, "%m/%d/%Y") - 30 )

prof.PrevStmt.fix <- temp
rm(temp)
```


Group user profile data by : Id, State
```{r}
temp <- prof.PrevStmt.fix %>%
  group_by(DebtDimId, State) %>%
  summarise(StartDay = min(as.Date(DatePrevStmt, "%m/%d/%Y" )),
          EndDay = max(as.Date(StatementDt, "%m/%d/%Y"))) %>%
  select(one_of(c('DebtDimId','State', 'StartDay', 'EndDay')))

length(unique(prof.dat$DebtDimId))

nrow(temp)

tbl.Member.HomeState <- temp
rm(temp, prof.PrevStmt.fix)


```

### 3. Table 2: DebtDimId, Restaurant name, Locate State, TransDate

All transaction with MCCcode: 5812 or 5814 
```{r}
temp <- trans.dat[which(trans.dat$MCCcode==5812 | trans.dat$MCCcode==5814),]

```

Remove transactions in which FDRMerchDescr contains phone number
```{r}
phone.reg.pattern = '\\s*(?:\\+?(\\d{1,3}))?[- (]*(\\d{3})[- )]*(\\d{3})[- ]*(\\d{4})(?: *[x/#]{1}(\\d+))?\\s*'
idx <- grep(phone.reg.pattern, temp$FDRMerchDescr)
if(length(idx) != 0)
{
  temp <- temp[-idx,]  
}

rm(phone.reg.pattern, idx)
```

Get State code from FDRMerchDescr
```{r}
state.reg.pattern = '[[:alpha:]]{2}$'
store.loc <- str_extract(temp$FDRMerchDescr, state.reg.pattern)
rm(state.reg.pattern)
```

Store Name 1: FDRMerchDescr - Remove city and state
```{r}
store.name <- temp$FDRMerchDescr
tbDelete.reg.pattern = '[[:space:]]*[[:alpha:]]{2,} [[:alpha:]]{2}$'
store.name <- gsub(tbDelete.reg.pattern, '', store.name)
```

Store Name 2: Remove string after two space (or more)
```{r}
tbDelete.reg.pattern = '[[:space:]]*\\s{2,}[[:print:]]*$'
store.name <- gsub(tbDelete.reg.pattern, '', store.name)
```

Store Name 3: Remove string after digit
```{r}
tbDelete.reg.pattern = '[[:space:]]*[[:alpha:]]*\\d+[[:print:]]*$'
store.name <- gsub(tbDelete.reg.pattern, '', store.name)
```

Store Name 4: Remove string after #
```{r}
tbDelete.reg.pattern = '[[:space:]]*#[[:print:]]*$'
store.name <- gsub(tbDelete.reg.pattern, '', store.name)
```

Store Name 5: Remove string after @
```{r}
tbDelete.reg.pattern = '[[:space:]]*@[[:print:]]*$'
store.name <- gsub(tbDelete.reg.pattern, '', store.name)
```

Store Name 6: Remove string after -
```{r}
tbDelete.reg.pattern = '[[:space:]]*-[[:print:]]*$'
store.name <- gsub(tbDelete.reg.pattern, '', store.name)
```

Bind member id, transaction day, store name, store loc
```{r}
member.id <- temp$DebtDimId
transaction.Day <- temp$TimeDimDt
tbl.clean.trans <- as.data.frame(cbind( cbind(member.id, transaction.Day), cbind(store.name, store.loc) ))

rm(member.id, transaction.Day, store.name, store.loc)
```


4. Fill Home state into transaction table
```{r}
HomeState <- NULL
for(i in 1:nrow(tbl.clean.trans)){
  transId <- tbl.clean.trans$transaction.Day[i]
  transDay <- as.Date(tbl.clean.trans$member.id[i], "%m/%d/%Y")
  HS <- tbl.Member.HomeState %>%
    filter(DebtDimId == transId & StartDay <= transDay & EndDay >= transDay) %>%
    select(State)
  HomeState <- c(HomeState, HS)
}




```
<!-- ### 2. Find all utility transaction -->

<!-- ```{r} -->
<!-- trans.dat.util <- trans.dat[which(trans.dat$MCCcode == 4900),] -->

<!-- nrow(trans.dat.util) -->
<!-- head(trans.dat.util) -->
<!-- summary(trans.dat.util) -->

<!-- member.list.util <- trans.dat.util$DebtDimId -->

<!-- ``` -->

<!-- ### 3. Figure out which state the user is living in -->

<!-- Here I suggest using fast food/restaurant to figure out where the member are living in.  -->

<!-- If we cannot find any fast food/restaurant record, we then move to durg store/pharmacies.  -->

<!-- The third one we will try to use (if above two cannot give us the answer), is gas station. (Note: there are 2 different MCCcode for fuel supply, one of them is auto fuel only: 5542, the other one is Fuel Dealers – Coal, Fuel Oil, Liquefied Petroleum, Wood: 5983) -->

<!-- If still not, we will report how many we are failed to be located. -->

<!-- Bias? Yes. If someone never eat outside if they could cook (at home), however, they eat fast food when they are traveling. Then, the address of this member will be miss classified. -->


<!-- #### restaurant and fast food having MCCcode: 5812 and 5814 -->
<!-- ```{r} -->
<!-- trans.dat.food <- trans.dat[which(trans.dat$MCCcode == 5812 | trans.dat$MCCcode == 5814 ),] -->

<!-- member.not.in.food <- member.list.util[which(!( member.list.util %in% trans.dat.food$DebtDimId) )] -->

<!-- length(member.not.in.food) -->
<!-- ``` -->

<!-- In 873 members, we have 224 member has no food record. -->

<!-- #### durg store or pharmacies MCCcode: 5912 -->
<!-- ```{r} -->
<!-- trans.dat.drug <- trans.dat[which(trans.dat$MCCcode == 5912 ),] -->

<!-- member.not.in.durg.either <- member.not.in.food[which(!( member.not.in.food %in% trans.dat.drug$DebtDimId) )] -->

<!-- length(member.not.in.durg.either) -->
<!-- ``` -->

<!-- In 224 unkown member, 170 has no durg record. -->

<!-- #### Automated Fuel Dispenser MCCcode: 5542 -->

<!-- ```{r} -->
<!-- trans.dat.fuel <- trans.dat[which(trans.dat$MCCcode == 5542 ),] -->

<!-- member.not.located <- member.not.in.durg.either[which(!( member.not.in.durg.either %in% trans.dat.fuel$DebtDimId) )] -->

<!-- length(member.not.located) -->
<!-- ``` -->

<!-- After 3 different tech, we still have 109 member cannot be located.  -->




