---
title: "Homework 2 Part 1. My own question and solution"
author: "Runan Yao"
date: "September 21, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```

# Question:
1. In each state, which utility company supplies most of our custom? 
2. If we want money back from utility companies, which one will be the best target?

Note:
1. Utilities – Electric, Gas, Water, Sanitary Fees has a unique MCCcode: 4900.
2. These utility bills often go with a 800 number and address. As we being told in class, the address is the company address, not the user address.
3. We also make an assumption that, the utility are paid together in one transaction. (This may not true, but it could simplify the question)


# Solution:

## 1. Load data from file:
```{r}
trans1 <- read.csv('Data/April Vintage Transactions page 1.csv', stringsAsFactors=FALSE)
trans2 <- read.csv('Data/April Vintage Transactions page 2.csv', stringsAsFactors=FALSE)
trans2$DebtDimId <- as.character(trans2$DebtDimId)
trans3 <- read.csv('Data/April Vintage Transactions page 3.csv', stringsAsFactors=FALSE)
trans3$DebtDimId <- as.character(trans3$DebtDimId)
trans4 <- read.csv('Data/April Vintage Transactions page 4.csv', stringsAsFactors=FALSE)
trans4$DebtDimId <- as.character(trans4$DebtDimId)

trans.dat <- bind_rows( bind_rows(trans1,trans2),bind_rows(trans3,trans4) )
nrow(trans.dat)
head(trans.dat)
summary(trans.dat)

```

## 2. Find all utility transaction

```{r}
trans.dat.util <- trans.dat[which(trans.dat$MCCcode == 4900),]

nrow(trans.dat.util)
head(trans.dat.util)
summary(trans.dat.util)

member.list.util <- trans.dat.util$DebtDimId

```

## 3. Figure out which state the user is living in
Here I suggest using fast food/restaurant to figure out where the member are living in. 

If we cannot find any fast food/restaurant record, we then move to durg store/pharmacies. 

The third one we will try to use (if above two cannot give us the answer), is gas station. (Note: there are 2 different MCCcode for fuel supply)

If still not, we will report how many we are failed to be located.

Bias? Yes. If someone never eat outside if they could cook (at home), however, they eat fast food when they are traveling. Then, the address of this member will be miss classified.


### restaurant and fast food having MCCcode: 5812 and 5814
```{r}
trans.dat.food <- trans.dat[which(trans.dat$MCCcode == 5812 | trans.dat$MCCcode == 5814 ),]

member.not.in.food <- member.list.util[which(!( member.list.util %in% trans.dat.food$DebtDimId) )]

length(member.not.in.food)
```

In 873 members, we have 224 member has no food record.
 
### durg store or pharmacies MCCcode: 